\documentclass[a4paper, 10pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[top=0.75in, bottom=0.75in, left=0.55in, right=0.55in]{geometry}

\title{Gurpreet Singh CV}
\author{Gurpreet Singh}
\date{19 February, 2017}

\begin{document}

% \maketitle

\begin{center}
    {\huge Gurpreet Singh} \\
    \vspace{5mm}
    
    B326 Hall 10, IIT Kanpur $ \cdot $ guggu@iitk.ac.in $ \cdot $ 9005528086

	\rule{\paperwidth}{0.4pt}

\end{center}

CREATIVE, FLEXIBLE, and able to adapt to changing priorities. I am HONEST, DEDICATED and a RESPONSIBLE person. I am KEEN and very WILLING to learn and develop new skills. I believe in approaching a problem DYNAMICALLY and PRACTICALLY. I have a keen interest in COMPETITIVE PROGRAMMING, and WEB DESIGNING and DEVELOPMENT. Other than this, I enjoy sketching, listening to music, watching anime, and well mostly sleeping.

\section*{Education}

	\begin{center}
	
		{\renewcommand{\arraystretch}{1.4}
		
		\begin{tabular}{ p{5cm}|p{5cm}|p{2cm}|p{3.5cm} }
		
		\hline
		\textbf{Degree} & \textbf{Institute} & \textbf{Year} & \textbf{CPI/Percentage} \\
		
		\hline
		B. Tech \newline (Computer Science and Engineering) & Indian Institute of Technology & 2015- & 8.8 \\
		
		\hline
		CLASS XII \newline (CBSE) & Delhi Public School, Kalyanpur & 2015 & 94.4 \\
		
		\hline
		CLASS X \newline (CBSE) & Delhi Public School, Kalyanpur & 2013 & 9.6 \\		
		\end{tabular}
		}
	
	\end{center}

\section*{Technical Skills}

	\begin{itemize}
		\item \textbf{Programming Languages: } C, C++, Python, Octave, R, Bash
		\item \textbf{Web Development: } PHP, Javascript, Node.js, MySQL, CSS/HTML
		\item \textbf{Development Tools: } MATLAB, Photoshop/Gimp
	\end{itemize}

\section*{Work Experience}

	\begin{itemize}
		\item
			\textbf{Content Curation Tool} (Intern) \\
			\emph{(Inmobi, 1st Dec 2015 - 15th Jan, 2016)} \\
			
			Worked in the OnO team in strategic advertising for better product discovery and user experience to develop a Curator Tool for an App to scrape and render information from the web. \\ \\ Designed the main interface and set up the main engine to scrape and render information from various websites.
		
	\end{itemize}

\end{document}
